The Error Estimate of Short-Range Force Calculation in the Inhomogeneous Molecular Systems

### 2012-07-25 shortRange.inhomo/alltext/pubilsh.pre/resubmit/main.tex
revise according to pre

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textsf{#1}}

@@

\section{Introduction}
Non-bonded short-range interactions are encountered in nearly every
molecular simulation.  A naive idea to calculate short-range
interactions is to explicitly calculate all pairwise
interactions. This results in a computational cost scaling $\mathcal
O(N^2)$ per time step, which becomes rapidly inefficient as the number
of particles $N$ grows. A better way is to introduce a cut-off
radius, outside of which all pairwise interactions are simply
neglected. In combination with the cell list and the neighbor list
algorithms \cite{frenkel02b}, the total computational cost of the
short-range interactions can be reduced to an acceptable level of
$\mathcal O(N)$.  The cut-off radius neglects all pairwise interactions
outside the cut-off radius, so it introduces an error in the interaction
calculation. Since the cut-offed system can be viewed as an approximation of
the original system, it is very important to test the
convergence of properties of interest as a function of the cut-off
radius.
% before deriving conclusions on the origianl system.

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$ for the standard Lennard-Jones 6-12 interaction.
Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long-range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogeneity of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
Instead for inhomogeneous systems, in which the density of a certain type of
atom may change within the length scale of a few molecules, 
and thus lack of uniformity, 
the standard molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{M_1}\v a_1 + 
  \frac{i_2}{M_2}\v a_2 + 
  \frac{i_3}{M_3}\v a_3,\qquad 0\leq i_\alpha< M_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $M_\alpha$ is the number of
divisions on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{m_1, m_2, m_3} & =
  m_1\v a_1^\ast +
  m_2\v a_2^\ast +
  m_3\v a_3^\ast, \qquad 0\leq m_\alpha<M_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{m_1=-\infty}^\infty\sum_{m_2=-\infty}^\infty\sum_{m_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{m_1=0}^{M_1}\sum_{m_2=0}^{M_2}\sum_{m_3=0}^{M_3}
  \hat{K}_{\textrm{homo}}(\v k'_{m_1,m_2,m_3})\:
  \hat{\rho}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means the Fourier transform. The prime on $\v k_{m_1,m_2,m_3}$ means
that the periodic image $m_\alpha - M_\alpha$ should use instead of
$m_\alpha$ when $m_\alpha \geq M_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{m_1,m_2,m_3})
  & =
  \frac{V}{M_1M_2M_3}
  \sum_{i_1=0}^{M_1}\sum_{i_2=0}^{M_2}\sum_{i_3=0}^{M_3}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $r_u$. {Perfect
convergence is achieved by using $r_u = 30\sigma$
for the dispersion energy $-4\epsilon (\sigma / r)^6$, where $\epsilon$
and $\sigma$ are the energy and length units.}
Since integral~\eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large uniform cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}.
{
  However, since the computational effort scales as the
  cube of the cut-off radius, a slightly larger  cut-off
  will notably increase the computational effort.
  For example, increasing the cut-off radius
  from $4\sigma$ to $5\sigma$ will cost
  twice the computational effort.}
Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.
 In order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$,
so the total computational expense of 
this method is at least $\mathcal O(N^2\log N)$, which is
prohibitive to the on-the-fly application.

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.
The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision.
% a slightly (at most $r_c^{\textrm{step}}$) larger
% cut-off radius
{
  Because if $r_c^{\textrm{step}}$ is sufficiently small,
  increasing the cut-off radius by no more than $r_c^{\textrm{step}}$
will not waste too much computational cost.
Here we recommend that $r_c^{\textrm{step}}$ should be no more than
0$.25\sigma$ that increases
the computational effort by 27\% at $r_c = 3.0\sigma$.}
Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

In most cases, the ensemble average of the error force
does not change very fast comparing with the typical time step used in a MD
simulation. Therefore, it is not necessary to calculate
the correction force every time step. Instead, for example,
we calculate it every 40 time steps in the liquid-vapor equilibrium
testing system (see Sec.~\ref{sec:tmp2.1}), which takes
no more than 10\% of the total computational cost.
If the long-range interaction calculations (which needs FFTs anyway)
are involved, this ratio should be even lower.
The idea of the long-range force correction is very similar to those
proposed by earlier works~\cite{guo1997long, mecke1997molecular,
  janecek2006long, shen2007comparative}. In those works, the density
profile is assumed to be inhomogeneous along one direction, and uniform
on the other two directions.
The
system is subdivided into slabs perpendicular to the inhomogeneous direction to
calculate the energy/force correction. The advantage of our force
correction is that it has \emph{no} uniformity restriction on the density
profile of the system, so can be used in a wider range of
inhomogeneous simulations.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the periodic box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  In the rest of this paper,
all quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literature~\cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \ref{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.  {We find that (not shown here)
  the bigger system leads to the same equilibrium
  quantities as the smaller system at the same cut-off radius, so
  it is reasonable to
  use it to produce the better converged data at $r_c^\ast=13$ and compare them with
  those obtained in the smaller system.
}

@@

\begin{figure}
  \centering
  \includegraphics[]{error.uniform.eps}
  \caption{(Color online)
    The RMS force error distribution of the 16000 Lennard-Jones
    particle system at $T^\ast=0.85$ with a uniform cut-off radius of
    $r_c^\ast = 7.5$. The green (light gray) squares present the real error. The
    solid blue (dark gray) line is the error estimate by \eqref{eqn:tmp5}. The
    dashed blue (dark gray) line
    is the homogeneity error contribution, while the
    dotted blue (dark gray) line
    is the inhomogeneity error contribution. The
    density distribution of the system is presented by red (middle gray)
    solid line
    for reference.  All properties are properly averaged on the $y$
    and $z$ direction, therefore, the profiles are plotted along the
    $x$ direction.  }
  \label{fig:tmp2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{rcut.and.error.eps}
  \caption{ (Color online)
    The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with adaptive
    cut-off radius. The control error $\mathcal E^\ast_{\textrm{C}}$
    is the same as the maximum error in the uniform cut-off simulation
    with $r_c^\ast=7.5$.  The resulting cut-off radius distribution is
    presented by the red (middle gray) line. All other notations are the same as
    Fig.~\ref{fig:tmp2}.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{fcorr.and.error.eps}
  \caption{ (Color online)
    The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with long-range
    force correction, i.e. eqn. \eqref{eqn:define-fcorr}. The cut-off
    radius is $r^\ast_c = 7.5$. The $x$ component of the correction
    force is presented by the solid red (middle gray) line.
    The green (light gray) squares
    present the real error. The solid blue (dark gray)
    line presents the estimated
    homogeneity error. }
  \label{fig:tmp4}
\end{figure}

>>>>

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textit{#1}}

@@

\section{Introduction}
Non-bonded short-range interactions are encountered in nearly every
molecular simulation.  A naive idea to calculate short-range
interactions is to explicitly calculate all pairwise
interactions. This results in a computational cost scaling $\mathit{O}(N^2)$ per time step, which becomes rapidly inefficient as the number
of particles $N$ grows. A better way is to introduce a cut-off
radius, outside of which all pairwise interactions are simply
neglected. In combination with the cell list and the neighbor list
algorithms \cite{frenkel02b}, the total computational cost of the
short-range interactions can be reduced to an acceptable level of
$\mathit{O}(N)$.  The cut-off radius neglects all pairwise interactions
outside the cut-off radius, so it introduces an error in the interaction
calculation. Since the cut-offed system can be viewed as an approximation of
the original system, it is very important to test the
convergence of properties of interest as a function of the cut-off
radius.
% before deriving conclusions on the origianl system.

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$ for the standard Lennard-Jones 6-12 interaction.
Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long-range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogeneity of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
Instead for inhomogeneous systems, in which the density of a certain type of
atom may change within the length scale of a few molecules, 
and thus lack of uniformity, 
the standard molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy (force) continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathit{O}(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathit{O}(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{M_1}\v a_1 + 
  \frac{i_2}{M_2}\v a_2 + 
  \frac{i_3}{M_3}\v a_3,\qquad 0\leq i_\alpha< M_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $M_\alpha$ is the number of
divisions on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{m_1, m_2, m_3} & =
  m_1\v a_1^\ast +
  m_2\v a_2^\ast +
  m_3\v a_3^\ast, \qquad 0\leq m_\alpha<M_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{m_1=-\infty}^\infty\sum_{m_2=-\infty}^\infty\sum_{m_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{m_1=0}^{M_1}\sum_{m_2=0}^{M_2}\sum_{m_3=0}^{M_3}
  \hat{K}_{\textrm{homo}}(\v k'_{m_1,m_2,m_3})\:
  \hat{\rho}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means the Fourier transform. The prime on $\v k_{m_1,m_2,m_3}$ means
that the periodic image $m_\alpha - M_\alpha$ should use instead of
$m_\alpha$ when $m_\alpha \geq M_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{m_1,m_2,m_3})
  & =
  \frac{V}{M_1M_2M_3}
  \sum_{i_1=0}^{M_1}\sum_{i_2=0}^{M_2}\sum_{i_3=0}^{M_3}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $r_u$. {Perfect
convergence is achieved by using $r_u = 30\sigma$
for the dispersion energy $-4\epsilon (\sigma / r)^6$, where $\epsilon$
and $\sigma$ are the energy and length units.}
Since integral~\eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large uniform cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}.
{
  However, since the computational effort scales as the
  cube of the cut-off radius, a slightly larger  cut-off
  will notably increase the computational effort.
  For example, increasing the cut-off radius
  from $4\sigma$ to $5\sigma$ will cost
  twice the computational effort.}
Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.
 In order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathit{O}(N \log N)$,
so the total computational expense of 
this method is at least $\mathit{O}(N^2\log N)$, which is
prohibitive to the on-the-fly application.

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathit{O}(MN\log N)$.
The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision.
% a slightly (at most $r_c^{\textrm{step}}$) larger
% cut-off radius
{
  Because if $r_c^{\textrm{step}}$ is sufficiently small,
  increasing the cut-off radius by no more than $r_c^{\textrm{step}}$
will not waste too much computational cost.
Here we recommend that $r_c^{\textrm{step}}$ should be no more than
0$.25\sigma$ that increases
the computational effort by 27\% at $r_c = 3.0\sigma$.}
Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

In most cases, the ensemble average of the error force
does not change very fast comparing with the typical time step used in a MD
simulation. Therefore, it is not necessary to calculate
the correction force every time step. Instead, for example,
we calculate it every 40 time steps in the liquid-vapor equilibrium
testing system (see Sec.~\ref{sec:tmp2.1}), which takes
no more than 10\% of the total computational cost.
If the long-range interaction calculations (which needs FFTs anyway)
are involved, this ratio should be even lower.
The idea of the long-range force correction is very similar to those
proposed by earlier works~\cite{guo1997long, mecke1997molecular,
  janecek2006long, shen2007comparative}. In those works, the density
profile is assumed to be inhomogeneous along one direction, and uniform
on the other two directions.
The
system is subdivided into slabs perpendicular to the inhomogeneous direction to
calculate the energy or force correction. The advantage of our force
correction is that it has \emph{no} uniformity restriction on the density
profile of the system, so can be used in a wider range of
inhomogeneous simulations.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid-vapor
densities, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the periodic box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  In the rest of this paper,
all quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literature~\cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \ref{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.  {We find that (not shown here)
  the bigger system leads to the same equilibrium
  quantities as the smaller system at the same cut-off radius, so
  it is reasonable to
  use it to produce the better converged data at $r_c^\ast=13$ and compare them with
  those obtained in the smaller system.
}

@@

\begin{figure}
  \centering
  \includegraphics[]{pre.error.uniform.eps}
  \caption{(Color online)
    The RMS force error distribution of the 16000 Lennard-Jones
    particle system at $T^\ast=0.85$ with a uniform cut-off radius of
    $r_c^\ast = 7.5$. The green (light gray) squares present the real error. The
    solid blue (dark gray) line is the error estimate by \eqref{eqn:tmp5}. The
    dashed blue (dark gray) line
    is the homogeneity error contribution, while the
    dotted blue (dark gray) line
    is the inhomogeneity error contribution. The
    density distribution of the system is presented by red (middle gray)
    solid line
    for reference.  All properties are properly averaged on the $y$
    and $z$ direction, therefore, the profiles are plotted along the
    $x$ direction.  }
  \label{fig:tmp2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{pre.rcut.and.error.eps}
  \caption{ (Color online)
    The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with adaptive
    cut-off radius. The control error $\mathcal E^\ast_{\textrm{C}}$
    is the same as the maximum error in the uniform cut-off simulation
    with $r_c^\ast=7.5$.  The resulting cut-off radius distribution is
    presented by the red (middle gray) line. All other notations are the same as
    Fig.~\ref{fig:tmp2}.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{pre.fcorr.and.error.eps}
  \caption{ (Color online)
    The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with long-range
    force correction, i.e. eqn. \eqref{eqn:define-fcorr}. The cut-off
    radius is $r^\ast_c = 7.5$. The $x$ component of the correction
    force is presented by the solid red (middle gray) line.
    The green (light gray) squares
    present the real error. The solid blue (dark gray)
    line presents the estimated
    homogeneity error. }
  \label{fig:tmp4}
\end{figure}
